---
title: "Reproducible Research Project1"
author: "Neal"
date: "October 28, 2017"
output: html_document
---

Personal activity monitoring device data analysis

Prepartion Loading necessary packages
```{r}
library(plyr)
library(dplyr)
library(lattice)
```


Step 1 Loading and preprocessing the data

```{r}
df <- read.csv("activity.csv")
df$date <- as.Date(df$date, format = "%Y-%m-%d")
```


Step 2 What is mean total number of steps per day

```{r}
# Calculate the total number of steps taken per day
df_sum_perday <-ddply(df, .(date), summarize, 
                         sum = sum(steps, na.rm = TRUE))
hist(df_sum_perday$sum, main = "Total Number of Steps Taken Per Day", 
     xlab = "Steps")
# Calculate and report the mean and median of the total number of steps taken per day
df_mean_perday <- ddply(df, .(date), summarize, 
                        mean = mean(steps, na.rm = TRUE))
df_mean_perday
df_median_perday <-ddply(df, .(date), summarize, 
                         median = median(steps, na.rm = TRUE))
df_median_perday

```

Step 3 What is the average daily activity pattern

```{r}
# Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
# and the average number of steps taken, averaged across all days (y-axis)
df_mean_interval <- ddply(df, .(interval), summarize,
                          mean = mean(steps, na.rm = TRUE))
plot(df_mean_interval$interval, df_mean_interval$mean, type = "l", 
     main = "Average number of steps taken per interval across all days", 
     xlab = "interval", ylab = "steps")
# Which 5-minute interval, on average across all the days in the dataset, 
# contains the maximum number of steps?
df_mean_interval$interval[which.max(df_mean_interval$mean)]
```

Step 4 Imputing missing values
```{r}
# Calculate and report the total number of missing values in the dataset
sum(is.na(df$steps))
# filling all the missing values, use the mean for that 5 minute interval
df_new <- left_join(df, df_mean_interval, by = c("interval" = "interval"))
df_new$steps[is.na(df_new$steps)] <- df_new$mean[is.na(df_new$steps)]
df_new <- select(df_new, steps:interval)
# Make a histogram of the total number of steps taken each day and Calculate 
# and report the mean and median total number of steps taken per day
df_new_sum_perday <-ddply(df_new, .(date), summarize, 
                         sum = sum(steps, na.rm = TRUE))
hist(df_new_sum_perday$sum, main = "Total Number of Steps Taken Per Day", 
     xlab = "Steps")
df_new_mean_perday <- ddply(df_new, .(date), summarize, 
                        mean = mean(steps, na.rm = TRUE))
df_new_mean_perday
df_new_median_perday <-ddply(df_new, .(date), summarize, 
                         median = median(steps, na.rm = TRUE))
df_new_median_perday

```

Step 5 Are there differences in activity patterns between weekdays and weekends?
```{r}
# Create a new factor variable in the dataset with two levels - "weekday" and 
# "weekend" indicating whether a given date is a weekday or weekend day.
df_new$dayType <- ifelse(weekdays(df_new$date, TRUE) %in% c("Sat", "Sun"), 
                         "weekend", "weekday")
df_new$dayType <- as.factor(df_new$dayType)
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute 
interval (x-axis) and the average number of steps taken, averaged across all 
weekday days or weekend days (y-axis).

```{r}
df_new_weekend <- ddply(filter(df_new, dayType == "weekend"), .(interval), summarize,
                          mean = mean(steps, na.rm = TRUE))
df_new_weekday <- ddply(filter(df_new, dayType == "weekday"), .(interval), summarize,
                          mean = mean(steps, na.rm = TRUE))
par(mfrow = c(2, 1))
plot(df_new_weekend$interval, df_new_weekend$mean, main = "weekend", 
     xlab = "Interval", ylab = "Number of steps", type = "l")
plot(df_new_weekday$interval, df_new_weekday$mean, main = "weekday",
     xlab = "Interval", ylab = "Number of steps", type = "l")
```

It seems that weekend will produce more steps during the day.









